---
title: "exercise-12"
output: html_document
date: "2024-04-19"
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

### Load in libraries
```{r}
library(tidyverse)
library(skimr)
library(naniar)
library(MASS)
library(MuMIn)
```

### Read in file
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"

d <- read_tsv(f)
```

### Exploratory analysis using skim from {skimr}
```{r}
skim(d)

### There are 4 categorical variables: order, family, Genus, and species
### There are 10 numeric variables: mass, gestation, newborn, weaning,
### weaning mass, AFR, max. life, litter size, litters/year, and refs
```

### "Wrangle" Dataset
```{r}
### Replace NAs, winnow dataset, and log transform numeric variables
d <- d %>%
  replace_with_na_all(condition = ~.x == -999.00) %>%
  subset(select=-c(refs,`litter size`)) %>%
  mutate( log_Mass = log(`mass(g)`),
          log_Gest = log(`gestation(mo)`),
          log_NewB = log(`newborn(g)`),
          log_WeanMo = log(`weaning(mo)`),
          log_WeanMass = log(`wean mass(g)`),
          log_AFR = log(`AFR(mo)`),
          log_MaxLife = log(`max. life(mo)`),
          log_LitpYear = log(`litters/year`))

### Create new "relative" variables by regressing log transformed variables
### on log transformed body mass. 
d <- d %>%
  mutate(relGest = resid(lm(log_Gest ~ log_Mass, data = d,
                            na.action=na.exclude)),
         relWean = resid(lm(log_WeanMo ~ log_Mass, data = d,
                            na.action=na.exclude)),
         relAFR = resid(lm(log_AFR ~ log_Mass, data = d,
                           na.action=na.exclude)),
         relLife = resid(lm(log_MaxLife ~ log_Mass, data = d,
                            na.action=na.exclude)),
         relNewbornMass = resid(lm(log_NewB ~ log_Mass, data = d,
                                   na.action=na.exclude)),
         relWeaningMass = resid(lm(log_WeanMass ~ log_Mass, data = d,
                                   na.action=na.exclude)),)
```

### Plot relative max life, newborn mass, and weaning mass by order
```{r}
### Relative max life by order plot

### Primates appear to have the highest residual lifespan followed by
### Xenartha and Scandentia 
rMLp <- ggplot(data = d, aes(x = order, y = relLife, fill = order)) +
  geom_boxplot() +
  xlab("Order") +
  ylab("Relative Max Life") +
  theme(axis.text.x = element_text(angle = 90))
rMLp

### Relative newborn mass by order plot

### Macroscelidea appear to have the highest residual newborn mass, though
### their data is limited, they are follwed by Cetacea and possibly
### Perissodactyla
rNMp <- ggplot(data = d, aes(x = order, y = relNewbornMass, fill = order)) +
  geom_boxplot() +
  xlab("Order") +
  ylab("Relative Newborn Mass") +
  theme(axis.text.x = element_text(angle = 90))
rNMp

### Relative newborn mass by order plot

### Perissodactyla appear to have the highest residual weaning mass, but ### again have little data. This is followed by Cetatea. 
rWMp <- ggplot(data = d, aes(x = order, y = relWeaningMass, fill = order)) +
  geom_boxplot() +
  xlab("Order") +
  ylab("Relative Weaning Mass") +
  theme(axis.text.x = element_text(angle = 90))
rWMp
```

### Run models and a model selection process of max life as a response
### variable with predictors: gestation, newborn, weaning, wean mass,
### litters/year, and overall body mass
```{r}
### Subset data for max life response variable and predictors
s <- d %>%
  subset(select = c(log_MaxLife, log_Gest, log_NewB, log_WeanMo, 
                             log_WeanMass, log_LitpYear, log_Mass)) %>%
  drop_na()

### Use dredge {MuMIn} to see model with lowest AICc
### the model with gestation, litters per year, mass and weaning as the
### predictors is the best model
MLm <- lm(log_MaxLife ~ log_Gest + log_NewB + log_WeanMo + log_WeanMass +
            log_LitpYear + log_Mass, data = s, na.action = na.fail) %>%
  dredge(beta = "none")

head(MLm)

### 5 models had a delta AICc < 4
### Gestation, Litters per Year, and Weaning were present in all of these
### models
MLm_avg <- summary(model.avg(MLm, subset = delta < 4, fit = TRUE))

MLm_avg

### Using a 95% confidence set the best model still has gestation, litters
### per year, mass, and weaningm, but there are now 9 "top" models
MLm_avg <- summary(model.avg(MLm, subset = cumsum(weight) <= 0.95, fit = TRUE))

MLm_avg

### Calculate 95% CI for model-averaged coefficients
confint(MLm_avg)

### Plot model-averaged coefficients and their CIs, across models
### gestation and litters per year remain significant, though mass
### and weaning are very close to zero.
plot(MLm_avg, full = TRUE, intercept = FALSE)
``` 

### Run models and a model selection process of AFR as a response
### variable with predictors: gestation, newborn, weaning, wean mass,
### litters/year, and overall body mass
```{r}
### Subset data for max life response variable and predictors
s <- d %>%
  subset(select = c(log_AFR, log_Gest, log_NewB, log_WeanMo, 
                             log_WeanMass, log_LitpYear, log_Mass)) %>%
  drop_na()

### Use dredge {MuMIn} to see model with lowest AICc
### the model with gestation, litters per year, mass, and weaning as the
### predictors is the best model
AFRm <- lm(log_AFR ~ log_Gest + log_NewB + log_WeanMo + log_WeanMass +
            log_LitpYear + log_Mass, data = s, na.action = na.fail) %>%
  dredge(beta = "none")

head(AFRm)

### 7 models had a delta AICc < 4
### Gestation, Litters per Year, and Weaning were present in all of these
### models
AFRm_avg <- summary(model.avg(AFRm, subset = delta < 4, fit = TRUE))

AFRm_avg

### Using a 95% confidence set the new "best model" has gestation, litters
### per year, mass, and weaning, with 9 models meeting this condition.
AFRm_avg <- summary(model.avg(AFRm, subset = cumsum(weight) <= 0.95, fit = TRUE))

AFRm_avg

### Calculate 95% CI for model-averaged coefficients
confint(AFRm_avg)

### Plot model-averaged coefficients and their CIs, across models only
### gestation and litters per year are significant
plot(AFRm_avg, full = TRUE, intercept = FALSE)
``` 
